---
title: "Work with your own data"
author: "Miquel De Cáceres/Javi de la Casa"
date: "2026-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Work with your own data {background-color="orange"}

## Exercise objectives

1.  Build trajectory objects using your own data
2.  Quantify and compare different trajectories
3.  Ask yourself questions about the fit of your data into ETA framework

-   If you did not bring your own data, you can use the complete lake database, that includes 5 lakes

## Step 1.

Load your data. Alternatively you can load the Azores data including three more lakes (exercises/StudentCSVdata/ETA_azores_exercise.csv)

```{r}

```


## Step 2.

Transform the data. how does your data look like? Transformation matters so give it a thought

```{r}

```


## Step 3.

Define the distance matrix (choose a distance metric), site and surveys

```{r}

```

## Step 4.

Create the trajectory object and plot it in a selected time slot. Describe visually how are they shifting, and if possible, in which direction (use loadings/ weights)

```{r}

```


## Step 5.

Calculate, using moving windows, the internal variance, speed and directionality

```{r}

```


## Step 6.

Compare the different trajectories using the RMA plot. Is it the same using all period and your selected period?

```{r}

```

